---
title: "Getting Started with persuasio"
author: "Xinrui Chen"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with persuasio}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## What is this for?

This vignette shows the basic usage of the `persuasio` package for estimation
and inference on average persuasion rates (APR) and local persuasion rates
(LPR) in the potential-outcomes framework of Jun and Lee (2023).

The package is designed for settings with a two-stage causal structure:

$$Z_i \text{ (instrument)} \rightarrow T_i \text{ (exposure)} \rightarrow Y_i \text{ (outcome)}$$

where $Z_i$ is a binary instrument, such as random assignment or policy
eligibility; $T_i$ is actual exposure to a persuasive message, such as media
consumption or campaign contact; and $Y_i$ is a binary behavioral outcome, such
as voting, donating, or purchasing. For an overview of empirical settings that
motivate persuasion estimands, see DellaVigna and Gentzkow (2010, *Annual
Review of Economics*,
[doi:10.1146/annurev.economics.102308.124309](https://doi.org/10.1146/annurev.economics.102308.124309)).

### When is `persuasio` applicable?

The framework applies whenever the empirical setting features:

- **A binary or discretized outcome**: vote/not vote, donate/not donate.
- **A persuasive channel**: media exposure, information shocks, campaigns, or
  policy eligibility.
- **A credible instrument**: randomized controlled trial assignment or a
  quasi-random rollout.
- **Imperfect compliance**: not everyone assigned receives treatment, or some
  untreated units still receive exposure.

The richest case uses individual-level data on $(Y, T, Z)$. Some estimands can
also be computed when only $(Y, Z)$ are observed, or when only summary
statistics are available.

### Key estimands

- **Average Persuasion Rate (APR)**: the population-wide persuasion rate. It is
  analogous to the average treatment effect (ATE), and is generally partially
  identified and reported as sharp bounds.
- **Local Persuasion Rate (LPR)**: the persuasion rate among compliers, the
  subpopulation for whom $Z_i$ changes $T_i$. It is analogous to the local
  average treatment effect (LATE) in the instrumental variables literature.

## `persuasio` TL;DR

`persuasio` is the recommended function for all estimation and inference. The
lower-level functions (`aprlb`, `aprub`, `lpr4ytz`, and `calc4persuasio`) are
available for direct access when only a specific component is needed.

### Choosing an estimator (`est`) in `persuasio`

| Situation                                      | `est` argument |
|------------------------------------------------|----------------|
| APR bounds + CI, have treatment variable       | `"apr"`        |
| LPR point estimate + CI                        | `"lpr"`        |
| APR bounds + CI, only outcome and instrument   | `"yz"`         |
| Bounds from summary statistics                 | `"calc"`       |

### Choosing an inference method in `persuasio`

| Situation                             | Recommended `method` |
|---------------------------------------|----------------------|
| No covariates, large sample           | `"normal"`           |
| Covariates present                    | `"bootstrap"`        |
| Small sample regardless of covariates | `"bootstrap"`        |

### Choosing a model in `persuasio` when covariates are present

| Identifying assumption                                 | `model`             |
|--------------------------------------------------------|---------------------|
| Homogeneous covariate effects across instrument groups | `"no_interaction"`  |
| Heterogeneous covariate effects across groups          | `"interaction"`     |
| Unsure                                                 | `"no_interaction"`  |

### Package setup

```{r setup}
library(persuasio)
```

The `GKB` data set is included in `persuasio` as lazy data, so it is available
after loading the package. For a full description of the dataset, see `GKB.R`
in the `R` folder or run `?GKB` in the console.

## `persuasio`: Unified Wrapper for Persuasion Effect Estimation

`persuasio` is the recommended entry point for the package. Rather than calling
`aprlb`, `aprub`, `lpr4ytz`, or `calc4persuasio` directly, `persuasio` provides
a single interface that handles input parsing and dispatches to the correct
estimator based on the `est` argument.

### Argument overview

| Argument  | Purpose                                                        | Default              |
|-----------|----------------------------------------------------------------|----------------------|
| `est`     | Which estimand to target (`"apr"`, `"lpr"`, `"yz"`, `"calc"`)  | Required             |
| `varlist` | Variable names in estimator-specific order                     | Required             |
| `data`    | Data frame                                                     | Required             |
| `level`   | Confidence level                                               | `0.95`               |
| `model`   | `"no_interaction"` or `"interaction"`                          | `"no_interaction"`   |
| `method`  | `"normal"` or `"bootstrap"`                                    | `"normal"`           |
| `nboot`   | Bootstrap replications, only when `method = "bootstrap"`       | `50`                 |

### `varlist` ordering

For `est = "apr"`, `"lpr"`, and `"calc"`, variables must be supplied in the
order `c(y, t, z, x, ...)`. For `est = "yz"`, where the treatment variable is
not observed, use `c(y, z, x, ...)`.

### Examples using the GKB dataset

**Example 1: Average persuasion rate (APR)**

```{r}
persuasio(
   est = "apr",
   y = "voteddem_all",
   t = "readsome",
   z = "post",
   level = 0.80,
   method = "normal",
   data = GKB
 )
```

This dispatches to `aprlb` and `aprub` to produce lower and upper bounds on the
APR with asymptotic normal confidence intervals at the 80% level.

**Example 2: Local persuasion rate (LPR)**

```{r}
persuasio(
   est = "lpr",
   y = "voteddem_all",
   t = "readsome",
   z = "post",
   level = 0.80,
   method = "normal",
   data = GKB
 )
```

This dispatches to `lpr4ytz` to estimate the persuasion effect among compliers
and returns standard errors and confidence intervals under the normal
approximation.

**Example 3: Outcome-instrument bounds with a covariate and bootstrap**

```{r}
persuasio(
  est = "yz",
  y = "voteddem_all",
  z = "post",
  x = "MZwave2",
  data = GKB,
  level = 0.80,
  model = "interaction",
  method = "bootstrap",
  nboot = 1000
)
```

The `"yz"` estimator uses only the outcome and instrument, making it suitable
when the treatment variable is not observed. `MZwave2` is passed as a covariate
with `model = "interaction"`, allowing heterogeneous covariate effects across
instrument groups. One hundred bootstrap replications are used here for
computational efficiency; at least 1,000 are recommended for applied research.

## Base estimators

This section details the lower-level functions in `persuasio`.

| Situation                                      | Base function       |
|------------------------------------------------|---------------------|
| Lower bound on APR only                        | `aprlb`             |
| Upper bound on APR only                        | `aprub`             |
| Local persuasion rate (LPR) for compliers      | `lpr4ytz`           |
| Bounds from published group means, no microdata | `calc4persuasio`    |

### `calc4persuasio`: Bounds from Summary Statistics

`calc4persuasio` computes lower and upper bounds on persuasion rates directly
from group means by instrument status. This function is useful when only
published summary statistics are available, or as a quick sanity check against
`aprlb` and `aprub`.

The function uses:

| Argument | Meaning                                            |
|----------|----------------------------------------------------|
| `y1`     | Mean of binary outcome `y` when instrument `z = 1` |
| `y0`     | Mean of binary outcome `y` when instrument `z = 0` |
| `e1`     | Mean of binary treatment `t` when instrument `z = 1` |
| `e0`     | Mean of binary treatment `t` when instrument `z = 0` |

#### Examples using the GKB dataset

Since `calc4persuasio` works on summary statistics, we first compute group
means manually:

```{r}
voteddem_all_0 <- mean(GKB$voteddem_all[GKB$post == 0], na.rm = TRUE)
voteddem_all_1 <- mean(GKB$voteddem_all[GKB$post == 1], na.rm = TRUE)

readsome_0 <- mean(GKB$readsome[GKB$post == 0], na.rm = TRUE)
readsome_1 <- mean(GKB$readsome[GKB$post == 1], na.rm = TRUE)
```

Then pass them to `calc4persuasio`:

```{r}
calc4persuasio(
  y1 = voteddem_all_1,
  y0 = voteddem_all_0,
  e1 = readsome_1,
  e0 = readsome_0
)
```

### `aprlb`: Lower Bound on the Average Persuasion Rate

`aprlb` estimates the lower bound on the average persuasion rate (APR)
following Jun and Lee (2023).

The function requires:

| Argument | Role                  | Must be binary? |
|----------|-----------------------|-----------------|
| `y`      | Outcome               | Yes             |
| `z`      | Instrument            | Yes             |
| `x`      | Covariates, optional  | No              |

#### Examples using the GKB dataset

**Example 1: No covariates**

```{r}
aprlb(y = "voteddem_all", z = "post", data = GKB)
```

This estimates the lower bound without covariates, returning a coefficient,
standard error, and 95% confidence interval.

**Example 2: With a covariate, no-interaction model**

```{r}
aprlb(y = "voteddem_all", z = "post", x = "MZwave2", data = GKB)
```

With covariates and the default `model = "no_interaction"`, the function fits a
single regression with the instrument and covariates. Analytical standard
errors are not reported in this case; use a bootstrap wrapper for inference.

**Example 3: By subgroup**

```{r}
gkb_groups <- split(GKB, GKB$MZwave2)
lapply(gkb_groups, function(sub_data) {
  aprlb(y = "voteddem_all", z = "post", data = sub_data)
})
```

This example computes `aprlb` separately for each value of `MZwave2` without
covariates, producing estimates with standard errors for each subgroup.

### `aprub`: Upper Bound on the Average Persuasion Rate

`aprub` estimates the upper bound on the average persuasion rate (APR) following
Jun and Lee (2023).

The function requires:

| Argument | Role                  | Must be binary? |
|----------|-----------------------|-----------------|
| `y`      | Outcome               | Yes             |
| `t`      | Treatment             | Yes             |
| `z`      | Instrument            | Yes             |
| `x`      | Covariates, optional  | No              |

#### Examples using the GKB dataset

**Example 1: No covariates**

```{r}
aprub(y = "voteddem_all", t = "readsome", z = "post", data = GKB)
```

Without covariates, `aprub` estimates the upper bound directly from the data
and returns a coefficient, standard error, and 95% confidence interval.

**Example 2: With a covariate, no-interaction model**

```{r}
aprub(y = "voteddem_all", t = "readsome", z = "post", x = "MZwave2", data = GKB)
```

With covariates and the default `model = "no_interaction"`, the function fits
single regressions with the instrument and covariates. Analytical standard
errors are not reported in this case; use a bootstrap wrapper for inference.

**Example 3: By subgroup**

```{r}
gkb_groups <- split(GKB, GKB$MZwave2)
lapply(gkb_groups, function(sub_data) {
  aprub(y = "voteddem_all", t = "readsome", z = "post", data = sub_data)
})
```

This example computes the no-covariate `aprub` estimates within each level of
`MZwave2`, returning estimates with standard errors for each subgroup.

### `lpr4ytz`: Local Persuasion Rate

`lpr4ytz` estimates the local persuasion rate (LPR), the average persuasion
effect among compliers. While `aprlb` and `aprub` bound the APR across the full
population, the LPR focuses on the complier subpopulation.

The function requires:

| Argument | Role                  | Must be binary? |
|----------|-----------------------|-----------------|
| `y`      | Outcome               | Yes             |
| `t`      | Treatment             | Yes             |
| `z`      | Instrument            | Yes             |
| `x`      | Covariates, optional  | No              |

#### Examples using the GKB dataset

**Example 1: No covariates**

```{r}
lpr4ytz(
  y = "voteddem_all",
  t = "readsome",
  z = "post",
  data = GKB
)
```

Without covariates, `lpr4ytz` returns the LPR estimate along with a standard
error and 95% confidence interval.

**Example 2: With covariate, no-interaction model**

```{r}
lpr4ytz(
  y = "voteddem_all",
  t = "readsome",
  z = "post",
  x = "MZwave2",
  data = GKB
)
```

The default `model = "no_interaction"` fits single regressions with the
instrument and covariates. This is a parsimonious specification and is
appropriate when there is no strong prior reason to expect heterogeneous
covariate effects.

**Example 3: With covariate, interaction model**

```{r}
lpr4ytz(
  y = "voteddem_all",
  t = "readsome",
  z = "post",
  x = "MZwave2",
  model = "no_interaction",
  data = GKB
)
```

The `model = "interaction"` specification fits separate models by instrument
group, allowing the effect of `MZwave2` to differ between the `z = 1` and
`z = 0` subgroups. Analytical standard errors are not reported in this case.

## Inference wrappers

Each inference wrapper calls the corresponding base function(s) for point
estimation, then adds confidence intervals via the chosen inference `method`:

- `persuasio4ytz` calls `aprlb` and `aprub` jointly, combining the two bounds
  into a single interval using either a Stoye (2009) normal correction or
  bootstrap resampling.
- `persuasio4yz` uses only the reduced-form relationship between outcome and
  instrument; the upper bound is fixed at 1.
- `persuasio4ytz2lpr` calls `lpr4ytz` for the LPR point estimate and wraps it
  with a standard normal or bootstrap confidence interval.

In all three cases, `persuasio` is the recommended entry point and dispatches to
the appropriate wrapper internally. Call the wrappers directly only when finer
control over arguments is needed.

### `persuasio4ytz`: Confidence Intervals for the Average Persuasion Rate

When `est = "apr"` in `persuasio`, `persuasio4ytz` is working under the hood. It
combines lower and upper bound estimates from `aprlb` and `aprub` into a single
confidence interval, using either an asymptotic normal approximation or
bootstrap resampling.

#### Examples using the GKB dataset

**Example 1: No covariates, asymptotic normal approximation**

```{r}
persuasio4ytz(
  y = "voteddem_all",
  t = "readsome",
  z = "post",
  method = "normal",
  level = 0.80,
  data = GKB
)
```

The asymptotic normal method is appropriate here because no covariates are
present and analytical standard errors are available from both `aprlb` and
`aprub`.

**Example 2: No covariates, bootstrap inference**

```{r}
persuasio4ytz(
  y = "voteddem_all",
  t = "readsome",
  z = "post",
  method = "bootstrap",
  level = 0.80,
  nboot = 1000,
  data = GKB
)
```

Bootstrap inference provides an alternative to the normal approximation; results
should be close to Example 1 in large samples. One hundred replications are used
here for computational efficiency; at least 1,000 are recommended for applied
research.

**Example 3: With a covariate, interaction model, bootstrap inference**

```{r}
persuasio4ytz(
  y = "voteddem_all",
  t = "readsome",
  z = "post",
  x = "MZwave2",
  model = "interaction",
  method = "bootstrap",
  level = 0.80,
  nboot = 1000,
  data = GKB
)
```

When covariates are present, bootstrap inference is recommended. Under
`model = "interaction"`, it is required because analytical standard errors are
not reported. The interaction specification allows heterogeneous covariate
effects across the `z = 1` and `z = 0` subgroups.

### `persuasio4ytz2lpr`: Confidence Intervals for the Local Persuasion Rate

When `est = "lpr"` in `persuasio`, `persuasio4ytz2lpr` is working under the
hood. It is the inference companion to `lpr4ytz`, wrapping the LPR point
estimate and adding confidence intervals via either a delta-method normal
approximation or bootstrap resampling.

With covariates, bootstrap inference is recommended. It is required when
analytical standard errors are unavailable, especially under interaction
models.

#### Examples using the GKB dataset

**Example 1: No covariates, normal inference**

```{r}
persuasio4ytz2lpr(
  y = "voteddem_all",
  t = "readsome",
  z = "post",
  method = "normal",
  level = 0.80,
  data = GKB
)
```

The function uses the delta-method standard error from `lpr4ytz` to construct an
80% confidence interval under the normal approximation.

**Example 2: No covariates, bootstrap inference**

```{r}
persuasio4ytz2lpr(
  y = "voteddem_all",
  t = "readsome",
  z = "post",
  method = "bootstrap",
  level = 0.80,
  nboot = 1000,
  data = GKB
)
```

Bootstrap inference provides an alternative to the normal approximation. One
hundred replications are used here for computational efficiency; at least 1,000
are recommended for applied research.

**Example 3: With a covariate, interaction model, bootstrap inference**

```{r}
persuasio4ytz2lpr(
  y = "voteddem_all",
  t = "readsome",
  z = "post",
  x = "MZwave2",
  model = "interaction",
  method = "bootstrap",
  level = 0.80,
  nboot = 1000,
  data = GKB
)
```

The `model = "interaction"` specification allows the effect of `MZwave2` to
differ between the `z = 1` and `z = 0` subgroups. Bootstrap inference is used
because analytical standard errors are not reported in this case.

### `persuasio4yz`: APR Bounds Without a Treatment Variable

When `est = "yz"` in `persuasio`, `persuasio4yz` is working under the hood. It
estimates bounds on the APR using only the binary outcome `y` and binary
instrument `z`, and is appropriate when only the outcome and instrument are
observed.

**Example 1: No covariates, normal inference**

```{r}
persuasio4yz(
  y = "voteddem_all",
  z = "post",
  method = "normal",
  level = 0.80,
  data = GKB
)
```

Without covariates, the lower-bound analytical standard error is available from
`aprlb`; the upper bound is fixed at 1. The normal approximation gives a
one-sided confidence interval for the identified set.

**Example 2: No covariates, bootstrap inference**

```{r}
persuasio4yz(
  y = "voteddem_all",
  z = "post",
  method = "bootstrap",
  level = 0.80,
  nboot = 1000,
  data = GKB
)
```

Bootstrap inference resamples the lower-bound estimator. One hundred
replications are used here for computational efficiency; at least 1,000 are
recommended for applied research.

**Example 3: With a covariate, interaction model, bootstrap inference**

```{r}
persuasio4yz(
  y = "voteddem_all",
  z = "post",
  x = "MZwave2",
  model = "interaction",
  method = "bootstrap",
  level = 0.80,
  nboot = 1000,
  data = GKB
)
```

With covariates, bootstrap inference is recommended. The
`model = "interaction"` specification allows heterogeneous covariate effects of
`MZwave2` across instrument groups.

## Related Software

The original Stata implementation is available at
<https://github.com/persuasio/persuasio-stata>.

## Reference

Jun, Sung Jae, and Sokbae Lee. 2023. "Identifying the Effect of Persuasion."
*Journal of Political Economy* 131 (8): 2032-2058.
<https://doi.org/10.1086/724114>.
